Trail Saving on Backtrack

### 2020-05-15 paper.tex
removed times font and cactus plot

\documentclass[runningheads]{llncs}
\usepackage{times}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
%\usepackage[noend]{algcompatible}
%\usepackage[ruled,linesnumbered,noend,oldcommands]{algorithm2e}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[table]{xcolor}
\usepackage{theorem}
\usepackage{float}

@@

\section{Experiments and Results}
We implemented our techniques in two different SAT solvers, MapleSAT
and Cadical, both of which have finished at or near the top of SAT
competitions for the past several years
\cite{Heule2018ProceedingsOS,Heule2019ProceedingsOS}. We then ran each
solver on the 800 total benchmark instances used in the main tracks of
the 2018 SAT Competition and 2019 SAT Race. The experiments were executed on a
cluster of 2.7 GHz Intel cores with 5000 seconds CPU time and 7 GB
memory limits for each instance. We chose not to output or verify the
proofs generated by any of the solvers. The Par-2 scores obtained and
total instances solved by each solver are reported in
Figures~\ref{fig:cadicalNew}, \ref{fig:cadicalOld}, and
\ref{fig:maple}. We also show the cactus plots of the new version of
cadical in Figure \ref{fig:cactus_cadical} and MapleLCMDist in Figure
\ref{fig:cactus_maple}.

In Figure~\ref{fig:cadicalNew} we used the newest version of cadical
(downloaded as of January 1, 2020) as the baseline solver, in
Figure~\ref{fig:cadicalOld} we used the version of cadical published
in \cite{DBLP:conf/sat/MohleB19} as the baseline, and in
Figure~\ref{fig:maple} we used MapleLCMDist \cite{DBLP:conf/ijcai/LuoLXML17,xiao2017maplelrb} as
the baseline. Each of the baselines were run with standard
non-chronological backtracking. We then refer to versions of each
solver with additional features implemented on top by adding
suffixes. "-chrono" refers to the solver with \cbt enabled (using the
solver's default settings), "-trail" refers to the baseline with plain
trail saving added (as described in Figure~\ref{fig:trailUse}),
"-trail-multipleBT" refers to the baseline with trail saving plus the
first enhancement of saving over multiple backtracks,
"-trail-multipleBT-lookahead" also adds the enhancement of lookahead
for conflicts by 2 decision levels, and
"-trail-multipleBT-lookahead-reason" also adds the final enhancement
to cease trail saving once a reason of "low quality" is reached. For
more details on the enhancements, please see section 4.2. The source
code for our implementation of trail saving can be obtained from the
web\footnote{https://github.com/rgh000/cadical-trail} or by downloading our submissions to the 2020 SAT Competition.

@@

\begin{figure}\centering\includegraphics[scale=0.65]{figures/cactus_maple_new.pdf}\caption{\small{Cactus plot for MapleLCMDist comparing standard non-chronological backtracking to \cbt and various configurations of trail saving. The first 300 problems were solved in less than 600 seconds, so that part of the plot is truncated.}}\label{fig:cactus_maple}\end{figure}

\section{Conclusion}
We have shown that our trail saving technique can speed up two
state-of-the-art \sat solvers, cadical and MapleSAT, as or more
effectively than chronological backtracking can without breaking the
conventional invariants of CDCL. We also introduced three enhancements
one can implement when using a saved trail and demonstrated
experimentally that these enhancements can sometimes improve a
solver's performance by a significant amount. We have shown that trail
saving and all enhancements we proposed are sound.

There are many avenues that can be pursued in future work, such as
using the saved trail to help make inprocessing techniques faster or
using the saved trail to learn multiple clauses from a single
conflict. It is also possible to combine trail saving with
chronological backtracking if one is willing to break the traditional
invariants of CDCL, but it would require further work to determine
whether or not this would be useful and how to best approach it.

>>>>

\documentclass[runningheads]{llncs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
%\usepackage[noend]{algcompatible}
%\usepackage[ruled,linesnumbered,noend,oldcommands]{algorithm2e}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[table]{xcolor}
\usepackage{theorem}
\usepackage{float}

@@

\section{Experiments and Results}
We implemented our techniques in two different SAT solvers, MapleSAT
and Cadical, \footnote{Our implementation is available at https://github.com/rgh000/cadical-trail.}
both of which have finished at or near the top of SAT
competitions for the past several years
\cite{Heule2018ProceedingsOS,Heule2019ProceedingsOS}. We then ran each
solver on the 800 total benchmark instances used in the main tracks of
the 2018 SAT Competition and 2019 SAT Race. The experiments were executed on a
cluster of 2.7 GHz Intel cores with 5000 seconds CPU time and 7 GB
memory limits for each instance. We chose not to output or verify the
proofs generated by any of the solvers. The Par-2 scores obtained and
total instances solved by each solver are reported in
Figures~\ref{fig:cadicalNew}, \ref{fig:cadicalOld}, and
\ref{fig:maple}. We also show the cactus plots of the new version of
cadical in Figure \ref{fig:cactus_cadical}.
%and MapleLCMDist in Figure \ref{fig:cactus_maple}.

In Figure~\ref{fig:cadicalNew} we used the newest version of cadical
(downloaded as of January 1, 2020) as the baseline solver, in
Figure~\ref{fig:cadicalOld} we used the version of cadical published
in \cite{DBLP:conf/sat/MohleB19} as the baseline, and in
Figure~\ref{fig:maple} we used MapleLCMDist \cite{DBLP:conf/ijcai/LuoLXML17,xiao2017maplelrb} as
the baseline. Each of the baselines were run with standard
non-chronological backtracking. We then refer to versions of each
solver with additional features implemented on top by adding
suffixes. "-chrono" refers to the solver with \cbt enabled (using the
solver's default settings), "-trail" refers to the baseline with plain
trail saving added (as described in Figure~\ref{fig:trailUse}),
"-trail-multipleBT" refers to the baseline with trail saving plus the
first enhancement of saving over multiple backtracks,
"-trail-multipleBT-lookahead" also adds the enhancement of lookahead
for conflicts by 2 decision levels, and
"-trail-multipleBT-lookahead-reason" also adds the final enhancement
to cease trail saving once a reason of "low quality" is reached. For
more details on the enhancements, please see section 4.2.

@@

%\begin{figure}\centering\includegraphics[scale=0.65]{figures/cactus_maple_new.pdf}\caption{\small{Cactus plot for MapleLCMDist %comparing standard non-chronological backtracking to \cbt and various configurations of trail saving. The first 300 problems were %solved in less than 600 seconds, so that part of the plot is truncated.}}\label{fig:cactus_maple}\end{figure}

\section{Conclusion}
We have shown that our trail saving technique can speed up two
state-of-the-art \sat solvers, cadical and MapleSAT, as or more
effectively than chronological backtracking can. We also introduced three enhancements
one can implement when using a saved trail and demonstrated
experimentally that these enhancements can sometimes improve a
solver's performance by a significant amount. We have shown that trail
saving and all enhancements we proposed are sound.

There are many avenues that can be pursued in future work, such as
using the saved trail to help make inprocessing techniques faster or
using the saved trail to learn multiple clauses from a single
conflict. It is also possible to combine trail saving with
chronological backtracking, but it would require further work to determine
whether or not this would be useful and how to best approach it.
